---
title: "System By Year"
output: html_document
params:
  cut1: 0
  cut2: 114
  cut3: 120
  cut4: 126
  cut5: 140
  lowpercent: 10
  startdate: !r as.Date("2016-01-01")
  enddate: !r as.Date("2017-01-01")
  href: "file:///C:/Users/Craig%20Nicholson/Documents/GitHub/SubstationVoltageAnalysis/"
  production: "J:/ElectSolve/Blue/Rcode/substationReport"
  localwin: "C:/Users/Craig Nicholson/Documents/GitHub/SubstationVoltageAnalysis/"
  localnix: "/Users/cn/rcode2_clean/Demo_20170403/"
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE, warning=FALSE, message=FALSE)
```

```{r Load Packages}
library(memisc)       # this clashes with dplyr rename, and select
                      # can i use something else
library(ggplot2)
library(ggExtra) # because remembering ggplot theme options is beyond me
library(viridis)
library(gridExtra)
library(dplyr)
library(tidyr)
library(RColorBrewer)
library(knitr)
library(lubridate)    # working with dates

```

#### Updated: `r format(Sys.Date(), format="%A, %B %d %Y")`  (`r Sys.time()`) 

```{r Functions}
# Trim leading and trailing whitespaces
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
```

```{r Set Working Directory}
setwd(params$localwin)
```

```{r Get the data via ODBC}
# Step 1, create ODBC on the box to the database, where R is installed.
#   this is for windows OS, and servers, and this can exist on the
#   application server.

# These only work on a windows box, and need to be uncommented when deployed
library(RODBC)       # need a way 
cnx <- odbcConnect("R",uid="electsolve",pwd="electsolve")
 
dtReadDayFrom <- params$startdate 
dtReadDayTo <- params$enddate
sSQLcmd <- paste(

"DECLARE @AMIReadSourceId INT
        SET @AMIReadSourceId = ( 
        SELECT ReadSourceId from 
           mdm.dbo.ReadSource 
           where ReadSourceDescription = 'ITRON') 

SELECT 
, h.meterreadintervalheaderid
, h.meteridentifier Meter
, h.ReadLogDate 'h.ReadDate'
, d.ReadDate 'i.ReadDate' 
, Readvalue Voltage
, l.locationNumber
, s.ScadaSubstationIdentifier Station

 FROM mdm.dbo.meterreadintervalheader h 
 INNER JOIN mdm.dbo.meter m
 ON m.meteridentifier = h.meteridentifier
 AND m.readsourceid = @AMIReadSourceId 
 INNER JOIN mdm.dbo.electricmeters em 
 ON em.meteridentifier= m.meteridentifier
 INNER JOIN mdm.dbo.MeterReadIntervalDetail d
 ON d.meterreadintervalheaderid = h.meterreadintervalheaderid
 AND d.ReadQualityCode = '3' --Good Reads Only
 INNER JOIN mdm.dbo.location l
 ON l.locationid = m.locationid
 INNER JOIN mdm.dbo.substation s
 ON s.substationid = l.substationid
 WHERE uom = 'voltage' AND readlogdate >= '",dtReadDayFrom," AND  readlogdate <= '",dtReadDayTo,"
 AND em.MeterVoltage like '240%'
 "
 ) 
") 
```

```{r Load the Data}
# Read date is in local time for both datasets CST or CDT
volts_raw <- sqlQuery(cnx, sSQLcmd)
##volts_raw <- read.csv("voltage_200420_2016.csv", header=TRUE) #38676 is from original query, v2 after delete is 38684
##volts_raw <- subset(volts_raw,volts_raw$Meter.ID == "200420")

```

```{r Remove Duplicates}
# First two tidy datasets for the raw data
# remove duplicates for voltage and power factor datasets
# had to add meter b/c the dupes were really not dupes just
# exact same reading on the data and for different meter
# and the meter data was not included in first output
# volts_clean <- volts_raw  %>% 
#   group_by(dtReadDate, Location, Substation, Voltage, Meter) %>% 
#   filter(row_number() == 1)
```

```{r Voltage - Remove all zero values of V - we are not concerned with outages}

# Remove all zero values, which indicates and outate
zeros <- nrow(subset(volts_raw, volts_raw$V <= 0))
volts_raw <- subset(volts_raw,volts_raw$V > 0)

```

```{r Voltage create new variable from 240v to 120v by dividing Voltage by 2}
# Divide all the votlages by 2 since we have 240v want to see 120v instead
# Save original value in new column to keep the original data idempotent, allowing us to re-run the same method multiple times.
volts_raw$V <- volts_raw$Voltage / 2

```

```{r Data Clean Up and Formatting}
# Gives us better column names
names(volts_raw)[1] <- "_id"
names(volts_raw)[2] <- "Meter"
names(volts_raw)[3] <- "ReadDate"
names(volts_raw)[3] <- "ReadDate"
names(volts_raw)[4] <- "Voltage"
names(volts_raw)[5] <- "Location"
names(volts_raw)[6] <- "Substation"

# Covert the chr date to POSIX
# Save original value in new column to keep the original data idempotent, allowing us to re-run the same method multiple times.
# This is the actual date of the interval reported from the endpoint
volts_raw$dtReadDate <- parse_date_time(volts_raw$ReadDate, orders="ymd HMS")

# Assumption all data is 15min intervals
# Ending Interval for all dates will have prefix (i), and i.ReadDate is the main date
# So when grouping use the i.dt's since they will be the day the intervals belong too
# instead of dtReadDate
# dtReadDate is the timestamp of the interval ending.
# 3/1/2017 00:00:00 really belongs to 2.28.2017
# See ending interval in Notes if you need further explanation
volts_raw$i.dtReadDate <-  volts_raw$dtReadDate - minutes(15)

#hm, used for reporting on the i.ReadDate for plots
volts_raw$hm.end <- format(as.POSIXct(volts_raw$i.dtReadDate, 
                           format="%Y-%m-%d %H:%M"),
                           format="%H:%M")
# Grouping and Rollups
# Add day and hour, so we can group by day or hour, using i.dtReadDate
# since ending interval will be the next day to keep consistency
# and for reviewing for issues
volts_raw$i.dtReadDay  <- floor_date(volts_raw$i.dtReadDate,"day")
volts_raw$i.year  <- year(volts_raw$i.dtReadDate)
volts_raw$i.month = month(volts_raw$i.dtReadDate, label=TRUE) #3 letter acroymn for US peeps
volts_raw$i.day  <- day(volts_raw$i.dtReadDate)
volts_raw$i.h <- hour(volts_raw$i.dtReadDate)
# WeekDay (M-F, Sat, Sun)
volts_raw$wday  <- wday(volts_raw$i.dtReadDay, label = TRUE)
volts_raw$Meter <- factor(volts_raw$Meter)

```

```{r Voltage Trim Substation Name}
# Trim and Upper the Substation Name for Consistency
volts_raw$Substation <- trim(volts_raw$Substation)

```

```{r Voltage Create Buckets}
# Create the buckets for the voltage
# This could be different based on each country.  
# These are US values from ANSI for 60Hz.
# (  0, 114],  <= 114
# (114, 126]   > 114 and <= 126
# (126, 140]   > 126 and <= 140
volts_raw$voltage.bucket <- cut(volts_raw$V,
                                    breaks = c(params$cut1,
                                               params$cut2,
                                               params$cut3,
                                               params$cut4,
                                               params$cut5))
                                               #labels=c('Low','Med', 'Best','High'))
```


```{r Scale the data between zero and one}

volts_raw$V.scale <- volts_raw$V/max(volts_raw$V)

```


```{r Log Returns}

# df2 <- volts_raw[,1:2]
# for(name in names(volts_raw)[2:length(names(volts_raw))]){
#     df2[,name] <- volts_raw[,name]
#     df2[2:nrow(df2),paste0(name, ".Log")] <- diff(log(as.numeric(as.character(volts_raw))), lag = 1)
# }




# clean up data (convert strings to numbers)
#df <- df %>% mutate_each(funs(as.numeric(as.character(.))), B126:H251)

# calculate log diff and merge
#df %>% merge(df %>% mutate_each(funs(c(NA,diff(log(.)))), B126:H251), by='Date', suffixes=c('','_log'))

# optionally apply rounding function
#df %>% mutate_each(funs(round(.,3)), B126_log:H251_log)

```

### Summary Statistics for - `r year(params$startdate)`
Substations : : `r length(unique(volts_raw$Substation))`
Meters collecting Voltage : `r length(unique(volts_raw$Meter))`  
Total Intervals collected : `r count(volts_raw)`   
Intervals with a voltage of zero removed : `r zeros`

```{r Summary of Voltage, eval=FALSE}
# Can I do this dynamically per month with group
summary(volts_raw$V)
```

#### Histogram - Distribution of Voltage

```{r Histogram, fig.height=8, fig.width=16}
      ggplot(data=volts_raw,aes(x=volts_raw$V)) +
        geom_histogram(binwidth = 0.5, color="black", fill="#333333") +
          facet_wrap(i.year~i.month, ncol=6)+
          theme_bw() +
                  ggtitle(paste("Meter ",params$meteridentifier, " | Histogram from",
                      format(count(volts_raw),big.mark=",",scientific=FALSE), 
                      "Intervals |", 
                      format(length(unique(volts_raw$Meter)),big.mark=",",scientific=FALSE),  
                      "Meters |",
                      length(unique(volts_raw$i.dtReadDay)), 
                      "day(s) | binwidth = 0.5"
                      ))

```



### Heatmap

```{r Plot heatmap of each individual meter in dataframe meters, fig.height=8, fig.width=16}
      volts_raw$hm.end <- factor(volts_raw$hm.end)
      volts_raw$interval<-as.numeric(volts_raw$hm.end)

      p <-ggplot(volts_raw,aes(i.day,interval,fill=V))+
      geom_tile(color= "white",size=0) +
      scale_fill_viridis(name=" Voltage",option ="C")
      p <-p + facet_grid(i.year~i.month) # This would work for each substation as well.

      #p <-p + scale_y_continuous(trans = "reverse", breaks = c(1,12,24,36,48,60,72,84,96))
      p <-p + scale_y_continuous( breaks = c(1,12,24,36,48,60,72,84,96))
      
      p <-p + scale_x_continuous(breaks =c(1,10,20,31))
      p <-p + theme_minimal(base_size = 16)
      p <-p + labs( x="Day", y="Hour Minute Commencing")
      p <-p + theme(legend.position = "bottom") +
      theme(plot.title=element_text(size = 18))+
      theme(axis.text.y=element_text(size=6)) +
      theme(strip.background = element_rect(colour="white"))+
      theme(plot.title=element_text(hjust=0))+
      theme(axis.ticks=element_blank())+
      theme(axis.text=element_text(size=7))+
      theme(legend.title=element_text(size=8))+
      theme(legend.text=element_text(size=6))+
      theme(panel.spacing.x=unit(0, "lines"),panel.spacing.y=unit(0.25, "lines")) +
                ggtitle(paste("Voltage for Meter -",params$meteridentifier, " | ",
                      format(count(volts_raw),big.mark=",",scientific=FALSE), 
                      "Intervals |", 
                      format(length(unique(volts_raw$Meter)),big.mark=",",scientific=FALSE),  
                      "Meters |",
                      length(unique(volts_raw$i.dtReadDay)), 
                      "day(s)"
                      ))+  
      removeGrid()#ggExtra
      p
      
```

### Load Shape Scaled Data
Scaled data is the current value / max value in the data.

```{r Load Shape, fig.height=8, fig.width=16}
      ggplot(data=volts_raw,aes(x=i.dtReadDate ,y=volts_raw$V.scale)) +
        geom_line(alpha = 1/1) +
        #scale_y_continuous(limits = c(110,130)) +
        facet_wrap(~i.year, ncol=1) +
        theme_bw() +
        ylab("Voltage Scaled") +
        xlab("Day") +
        theme(legend.position="none")  +
        #scale_x_discrete(breaks=c("00:00","04:00","08:00","12:00","16:00","20:00")) +
        ggtitle(paste("Meter ",params$meteridentifier, " | Load Curves from",
                      format(count(volts_raw),big.mark=",",scientific=FALSE), 
                      "Intervals |", 
                      format(length(unique(volts_raw$Meter)),big.mark=",",scientific=FALSE),  
                      "Meters |",
                      length(unique(volts_raw$i.dtReadDay)), 
                      "day(s)"
                      ))


```


# Data

```{r Voltage Summaries for Meters By Month}

v_yearmonthmeter <- volts_raw %>%
  group_by(Substation, Location, Meter, i.year, i.month, voltage.bucket)  %>%
  summarise( voltage.min = min(V),
             voltage.mean = mean(V),
             voltage.median = median(as.numeric(V)),
             voltage.max = max(V), 
             v.intervals = n())  %>% 
  arrange(Substation, Location, Meter, i.year, i.month, voltage.bucket)


```

```{r Summary Year Month By Substation}
# Review the total counts for voltage in each substation this goes with scatter plots 
# in bivar. or multivariate data

v_ym_results <-
  v_yearmonthmeter %>% 
    group_by(Substation, i.year, i.month, voltage.bucket) %>% 
    summarise(count = sum(v.intervals))  %>% 
        arrange(Substation, i.year, i.month, voltage.bucket)

# pivot the data for easy review
v_ym_results <- spread(v_ym_results, voltage.bucket, count)
# Replace .na values here
v_ym_results[is.na(v_ym_results)] <- 0



# what to do when the columns are missing or null
if(!"(0,114]" %in% colnames(v_ym_results))
{
  v_ym_results$"(0,114]" <- NA 
}

# what to do when the columns are missing or null
if(!"(126,140]" %in% colnames(v_ym_results))
{
  v_ym_results$"(0,114]" <- NA 
}

v_ym_results$total <- 
  v_ym_results$`(0,114]` +
  v_ym_results$`(114,120]` +
  v_ym_results$`(120,126]` +
  v_ym_results$`(126,140]`

# calculate the percent for each bucket
v_ym_results$Low <- 
  100*(v_ym_results$`(0,114]`/v_ym_results$total)

v_ym_results$Mid <- 
  100*(v_ym_results$`(114,120]`/v_ym_results$total)

v_ym_results$Nominal <- 
  100*(v_ym_results$`(120,126]`/v_ym_results$total)

v_ym_results$High <- 
  100*(v_ym_results$`(126,140]`/v_ym_results$total)
```


```{r Add link to sub reports}

v_ym_results$View <- paste0("[View]","(",
                            params$href,
                            v_ym_results$Substation,
                            v_ym_results$i.year,
                            v_ym_results$i.month,
                            ".html",
                            ")")

v_ym_results$View <- paste0("[View]", "(", params$href ,"SubstationByMonth.html)")

```

Sorting by Substation to help ease the manual review of the data.  Low, Mid, Nominal and High values are the percentages of intervals in this range.

```{r Voltage Sub Year Month with Low percentanges above zero}
kable(v_ym_results[order(v_ym_results$i.month),])
```

# Notes

* Reading with a voltage of zero {0} will be removed.
* Duplicate removal is turned off

## Ending Intervals
Electric industy practices use ending interval notation.  An example is provided below in 15min intervals:

Interval Date    | Value            | Interval Belongs to Day
-----------------|------------------|-------------------
**01/01/2000 00:00** | 12/31/1999 23:45 | **12/31/1999**
01/01/2000 00:15 | 01/01/2000 00:00 | 01/01/2000
01/01/2000 00:30 | 01/01/2000 00:15 | 01/01/2000
01/01/2000 00:45 | 01/01/2000 00:30 | 01/01/2000
01/01/2000 01:00 | 01/01/2000 00:45 | 01/01/2000
01/01/2000 01:15 | 01/01/2000 01:00 | 01/01/2000
01/01/2000 01:30 | 01/01/2000 01:15 | 01/01/2000
01/01/2000 01:45 | 01/01/2000 01:30 | 01/01/2000
01/01/2000 02:00 | 01/01/2000 01:45 | 01/01/2000
...              |                  |
01/01/2000 21:15 | 01/01/2000 21:00 | 01/01/2000
01/01/2000 21:30 | 01/01/2000 21:15 | 01/01/2000
01/01/2000 21:45 | 01/01/2000 21:30 | 01/01/2000
01/01/2000 22:00 | 01/01/2000 21:45 | 01/01/2000
01/01/2000 22:15 | 01/01/2000 22:00 | 01/01/2000
01/01/2000 22:30 | 01/01/2000 22:15 | 01/01/2000
01/01/2000 22:45 | 01/01/2000 22:30 | 01/01/2000
01/01/2000 23:00 | 01/01/2000 22:45 | 01/01/2000
01/01/2000 23:15 | 01/01/2000 23:00 | 01/01/2000
01/01/2000 23:30 | 01/01/2000 23:15 | 01/01/2000
01/01/2000 23:45 | 01/01/2000 23:30 | 01/01/2000
**01/02/2000 00:00** | 01/01/2000 23:45 | **01/01/2000**

## Bucket Explanations

Description | Value       | Representation
------------|-------------|-------------------
Low         |  (0,114]    | >   0 and <= 114  
Mid         |  (114,120]  | > 114 and <= 120  
Nominal     |  (120,126]  | > 120 and <= 126  
High        |  (126,140]  | > 126 and <= 140 

## Configuration
Pass in the workingdir as "J:/ElectSolve/Blue/Rcode", or where the directory is located.
This can be done from the r script which creates the report. 

Depending on the install location for the mark down file (.Rmd) the working directoty should be something 
similar to setwd("J:/ElectSolve/Blue/Rcode").   

## Parameters

#### Cut variables
cut1 : `r params$cut1`  
cut2 : `r params$cut2`  
cut3 : `r params$cut3`  
cut4 : `r params$cut4`   
cut5 : `r params$cut5`  
Used to slice the data in different categories.

#### Lower Percent
This variable is used to reduce the number of meters we return when viewing the result in plot Comparisons of Meters By Month 

lowpercent: `r params$lowpercent` 

#### Set Working Direcory

production: `r params$production`  
localwin: `r params$localwin`  
localnix: `r params$localnix`  

### Query
`r sSQLcmd`

## References
* Use Value current/Max(Value) for smoothing.
* Use Log Returns to normalize the data. eg.. log(Today's Value/ Yesterday's Value).
* Do we need auto-correlation to see if yesterday's data has an affect on today's data.
* Scatter Matrix, do we have correlation between meters, meter which behave the same? 
* Do meters have specific behaviors we can flag day to day before a trend develops, a trend we want to avoid, one that occurred historically?
* What about meters dropping intervals... what percentage of meters not reporting in is bad?
* What percentage of zeros is bad or affects the results?
* We have a handful of variables maybe there should be a weighted formula to figure out who is the worst performer.
  * Temperature
  * Voltage
  * Time of Day
  * Location (Hot arid, by ocean {salty air}, humidity, etc ...)
  * Line Conditions
* [Code Repository](https://electsolve.kilnhg.com/Code/RCode/Group/SubstationReport)